---
title: "Bellabeat Analysis"
author: "Oscar"
date: "2023-04-10"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Business Task
Analyze Fitbit user trends to find opportunities for Bellabeat's growth by influencing Bellabeat's marketing strategies.
<br> 

### Data Preparaion
In this phase, data is downloaded, imported, and prepped to be analyzed.

<br> 

#### Downloading Data
We will be using Fitbit users' data on Kaggle, as requested by Bellabeat. This data set tracked 30 users' daily, hourly, and minute activities / heart rate / steps.
<br> 


#### Loading Packages
```{r}
install.packages("tidyverse", repos = "http://cran.us.r-project.org")
library(tidyverse)
```
<br>

### Data Processing / Exploration
```{r}
activity <- read.csv("Fitabase Data 4.12.16-5.12.16/dailyActivity_merged.csv")
calories <- read.csv("Fitabase Data 4.12.16-5.12.16/dailyCalories_merged.csv")
intensities <- read.csv("Fitabase Data 4.12.16-5.12.16/hourlyIntensities_merged.csv")
sleep <- read.csv("Fitabase Data 4.12.16-5.12.16/sleepDay_merged.csv")


head(activity)
head(calories)
head(intensities)
head(sleep)

str(activity)
str(calories)
str(intensities)
str(sleep)

```

#### Data Manipulation: Cleaning and aggregating to find the average hourly intensity of users

Since the original format was in a string format, I had to convert the string to datetime in order to extract the HMS by reconverting to HMS in string format
```{r}
intensities$ActivityHour <- mdy_hms(intensities$ActivityHour)
intensities$ActivityHour = format(intensities$ActivityHour, "%H:%M:%S")


class(intensities$ActivityHour)
```
<br>

## Data Analysis


```{r}
n_distinct(activity$Id)
n_distinct(calories$Id)
n_distinct(intensities$Id)
n_distinct(sleep$Id)
```
With the 'n_distinct' function, we are able to identify that there are 33 participants in daily activities, calories, and intensities, while sleep has 24. Since the discrepancies between 33 and 24 is not too large, we will continue using these data sets for our analysis.

```{r}
activity %>% 
  select(VeryActiveMinutes, FairlyActiveMinutes, LightlyActiveMinutes, SedentaryMinutes, Calories) %>% 
  summary()
```
Looking at the daily activity's averages(mean) in this summary, we can see that the Sedentary Minutes(991 minutes or 16.5 hours) is more than 4x the combined minutes from Very Active, Fairly Active, and Lightly Active(combined sum: 228 minutes or 4 hours). This massive difference may give us a clue that most Fitbit users are inactive. Looking into the Sleep Minutes may give us a better clue as to how much of those Sedentary Minutes are spent asleep. 

```{r}
sleep %>% 
  select(TotalMinutesAsleep, TotalTimeInBed) %>% 
  summary()
```
It looks like users spend a little under half their inactivity in bed. Which still leaves 532 minutes or about 9 hours inactive outside of bed. With this 9 hour inactivity remainder, we can conclude that users may be busily inactive during their full time job, or their routine is full of inactivity.

We can get a better understanding of users' routine by looking at their average hourly intensity. From this table we can see that users are mostly active after 5 PM.

```{r}
avgIntensities <- intensities %>% 
  group_by(ActivityHour) %>% 
  summarize(TotalAvgIntensity = mean(TotalIntensity))

intensities %>% 
  group_by(ActivityHour) %>% 
  summarize(TotalAvgIntensity = mean(TotalIntensity))
```
<br>

## Data Visualization
```{r}
selectActivity <- activity %>% 
  select(VeryActiveMinutes, FairlyActiveMinutes, LightlyActiveMinutes, SedentaryMinutes)

activityPlot <- colMeans(selectActivity)

hours <- round(activityPlot / 60, digits = 1)
lbls <- paste(names(activityPlot), hours, "hours")

pie(activityPlot, col = c("#ffa600", "#ed5562", "#8d4280", "#1e345c"), labels = lbls)
```

Here we can see the massive coverage of Sedentary Minutes compared to all the other active states combined. From the plot, the Sedentary minutes take up approximately 16.5 hours or 81% of the total average activity. 

To get a clearer understanding, we can see when users are mostly inactive.
Here we can see the difference of intensities after 5PM from users. 

```{r}
ggplot(avgIntensities, aes(x=ActivityHour, y=TotalAvgIntensity)) +
  geom_bar(stat = "identity", aes(fill=TotalAvgIntensity)) +
  theme(axis.text.x = element_text(angle = 45)) +
  labs(title = "Average Hourly Intensity",
       x = "Hour",
       y = "Total Average Intensity")
```



## Conlusions & Reccommendations 

