---
title: "Bellabeat Analysis"
author: "Oscar"
date: "2023-04-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Business Task
Analyze Fitbit user trends to find opportunities for Bellabeat's growth by influencing Bellabeat's marketing strategies.
<br> 

### Data Preparaion
In this phase, data is downloaded, imported, and prepped to be analyzed.

<br> 

#### Downloading Data
We will be using Fitbit users' data on Kaggle, as requested by Bellabeat. This data set tracked 30 users' daily, hourly, and minute activities / heart rate / steps.
<br> 


#### Loading Packages
```{r}
install.packages("tidyverse", repos = "http://cran.us.r-project.org")
library(tidyverse)
```
<br>

### Data Processing / Exploration
```{r}
activity <- read.csv("Fitabase Data 4.12.16-5.12.16/dailyActivity_merged.csv")
calories <- read.csv("Fitabase Data 4.12.16-5.12.16/dailyCalories_merged.csv")
intensities <- read.csv("Fitabase Data 4.12.16-5.12.16/dailyIntensities_merged.csv")
sleep <- read.csv("Fitabase Data 4.12.16-5.12.16/sleepDay_merged.csv")


head(activity)
head(calories)
head(intensities)
head(sleep)

str(activity)
str(calories)
str(intensities)
str(sleep)

```



## Data Analysis
```{r}
n_distinct(activity$Id)
```



## Data Visualization

